---
title: "da-bivariate-climate-maps"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{da-bivariate-climate-maps}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Inputs
```{r setup}
library(surveyGEER)
library(rgee)
library(tidyverse)
library(sf)
library(tidyrgee)
devtools::load_all()
ee_Initialize()


con <- DBI::dbConnect(RPostgres::Postgres(),
                      dbname = "global_gdb",
                      user      = rstudioapi::askForPassword("Database user"),
                      password      = rstudioapi::askForPassword("Database password"),
                      port     = 5432)

# DBI::dbListTables(con)
admin1 <- st_read(con,"eth_adm1")
admin1 <- admin1 |> 
  rename(
    geometry="geom"
  )



chirps_link <- "UCSB-CHG/CHIRPS/DAILY"
chirps <- ee$ImageCollection(chirps_link)
chirps_tidy <- as_tidyee(chirps)


country_code <- "eth"
AOI_layer_name <- "eth_adm0"
AOI_grid_layer_name <- NULL
hex_diameter <- 15000 # 15 km set as de
surveyGEER:::aoi_or_grid_warning()

```


Create grid
```{r}

grid <- load_aoi_grid(aoi =AOI_layer_name,
                           grid = AOI_grid_layer_name,
                           hex_diameter = hex_diameter,
                           country_code = country_code
                           )

```
# CHIRPS monthly standard score
## ImageCollection maninpulation

- calculate time frame of interest precip values
- calculate baseline time frame comparable precip values
- join on common term


```{r}
rainfall_current <-  chirps_tidy |> 
  filter(year==2022) |> 
  group_by(year,month) |> 
  summarise(
    stat="sum"
  )

rainfall_yrmo_historical <-  chirps_tidy |> 
  filter(year %in% 1981:2020) |> 
  group_by(year,month) |> 
  summarise(
    stat="sum"
  ) 
rainfall_mo_historical <- rainfall_yrmo_historical |> 
  group_by(month) |> 
  summarise(stat=list("mean","sd"))


rainfall_comparison_historical <- rainfall_current |> 
  inner_join(rainfall_mo_historical, by="month")
```

from joined term we can calculate standard (Z) score. `$expression` methods have not yet been written into `tidyee` . Therefore we run conversion `as_ee` function to convert `tidyee` class back to `ImageCollection` class native to `rgee` to use `$expression` building syntax to create the standard score. After creation we can throw it back into our `tidyee` workflow with `as_ee`

```{r}
rainfall_comparison_historical_ic <- rainfall_comparison_historical |> 
  as_ee()
rainfall_comparison_historical_ic$first()$bandNames()$getInfo()

rainfall_comparison_historical <- rainfall_comparison_historical |> 
  select(precip_current="precipitation_sum",
         precip_baseline_mean="precipitation_sum_mean",
         precip_baseline_sd = "precipitation_sum_stdDev"
         )
rainfall_comparison_historical_ic <- rainfall_comparison_historical |> 
  as_ee()
rainfall_comparison_historical_ic$first()$bandNames()$getInfo()
chirps_zscore_ic<- rainfall_comparison_historical_ic$map(
  function(img){
    zscore<- img$expression(
      "float((precip_current-precip_baseline_mean)/(precip_baseline_sd))",
      opt_map= list(precip_current= img$select("precip_current"),
                    precip_baseline_mean= img$select("precip_baseline_mean"),
                    precip_baseline_sd= img$select("precip_baseline_sd")
      )
    )$rename("precip_z_score")
    img$select("precip_current","precip_baseline_mean")$addBands(zscore)
  }
  
) 


chirps_z_tidy <- as_tidyee(chirps_zscore_ic)
```

## Zonal Stats
Now time for the zonal 

```{r}

chirps_zonal <- chirps_z_tidy |> 
  ee_extract_tidy(
    y = grid,stat = "mean",scale = 5500,via = "drive"
                  )

```
