---
title: "function_building"
author: "Zack Arno"
date: "4/8/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```








Again I believe there should be 0/very low risk associated with this file on it's own. 



- you will need to choose a single date for your assessment. This should be a date during which the data collection took place. The best dates to choose are either a.) start of assessment, b.) end of assessment, c.) middle of assessment
- The assessment date object created here will be used in later `ee_` functions
```{r}


# choose assessment date 
# assessment_date <- 

assessment_date <-  median_survey_date(df = dat, survey_date = "survey_date")







```


## Data setup
This creates a new geodata file (`gd.rds`) in the `dat` folder. It can be linked back to the original assessment data with `lt.rds` in the `dat` folder.

This should only really run this once. In fact, you will get an error if you run it again.
```{r}
# debugonce(setup_geo_data)
setup_geo_data(df = dat, 
                uuid = "_uuid",
                lon = "_gps_reading_longitude",
                lat = "_gps_reading_latitude")
```

overwrite the dat object with the new anonymized geodata
```{r}
dat <- read_rds(file = here::here("dat/gd.rds"))
lt <- read_rds(file = here::here("dat/lt.rds"))
path_secret <- file.path("dat", "lt2.rds")


k <- openssl::aes_keygen()
key <- cyphr::key_openssl(k)

pw_encrypted <- cyphr::encrypt_string("buttercuppeanut", key)
cyphr::decrypt_string(pw_encrypted, key)

cyphr::encrypt_object(object = lt, key = key, path_secret)

bla <- cyphr::ssh_keygen(path = "dat",password = "butternut")







sec_key <- cyphr::key_openssl(aes("mysecretkey"))

key


cyphr::encrypt_object(object = lt, key = key, path_secret)

key
read_rds("dat/lt_secret.rds")

k <- openssl::aes_keygen()

secret <- cyphr::encrypt_string("my secret string", key)
cyphr::decrypt_string(secret, key)

path_secret <- "dat/lt2.rds"
cyphr::encrypt_object(obj, key, path_secret)

```

now we are ready for extraction

# how about NDVI

What date range do we want to extract modis for? Recomment

- a monthly composite composed of dates 1 month prior to assessment
- how to deal with clouds?
- need to set some contextual parameters (these should go at top)



```{r}
final_month_of_last_grow_season <-  "2021-09-01"

date_range <- date_to_range(date = assessment_date,make_range = "backward",by = 1,unit = "month")

library(rgee)
ee_Initialize()

#' extract NDVI deviance (z scores)
#' @param dat sf object
#' @param date assessment date to use
#' @param ic_url `ee$ImageCollection` source url
#' @description provide point data, date, and GEE multi-band imagery URL and
#' function will calculate the monthly NDVI composite for the date provided, the three previous months and compare that to 
#' a historical baseline
#' @return data.frame containing original sf object identified column along with the NDVI for the assessment month and previous 3 month Z-scores 
#' compared to the historical baseline for those months.
#' 
#' 
#' @examples \dontrun{
#' # need to finish example
#' # dat =  load_dat()
#'  assessment_date <- "2021-06-014"
#'  extract_ndvi_z(dat=dat, date=assessment_date, ic_url="MODIS/006/MOD13Q1")
#' 
#' }
#' 
#' 




extract_ndvi_z <-  function(dat=dat,date=assessment_date,ic_url="MODIS/006/MOD13Q1"){
  
  
  run_date <- Sys.Date()
  # v0.1 should support MODIS and hopefully l8, need to figure out if we can make historical baseline NDVI for S2 (linear regression?)
  # assertthat::assert_that(ic_url %in% c(modis,sentinel,l8))
  cat("creating bounding bbox_ee\n")
  bbox_ee<- rgee::sf_as_ee(sf::st_bbox(dat) |> sf::st_as_sfc())
  ic <- rgee::ee$ImageCollection(ic_url)$filterBounds(bbox_ee)
  if(ic_url=="MODIS/006/MOD13Q1"){
    baseline_years <- c(2000,2015) # may want to parameterize this
    ic <- ic$select("NDVI")
    scale <- 250
    src_suffix <- "MODIS"
  }
    ic_dates <- rgee::ee_get_date_ic(ic)
    last_img_meta<- ic_dates |> 
      mutate(
        time_start = lubridate::ymd(time_start),
        days_since= time_start-lubridate::ymd(assessment_date)#,
        # days_since_abs= abs(lubridate::ymd(time_start)-lubridate::ymd(assessment_date))
      ) |>
      # MVP start by looking backwards only - can develop algorithm to decide when okay/better to look forward
      filter(days_since<=0) |> 
      slice_max(order_by = days_since,n = 1) 
    
    last_img_month = lubridate::month(last_img_meta$time_start)
    mr<- c(last_img_month-3,last_img_month)
  
    
    
    moi_mean_composites <- easyrgee::ic_yr_mo_composite_stat(ic = ic,
                                      month_range = c(mr[1],mr[2]),
                                      year_range = c(baseline_years[1],baseline_years[2]),
                                      stat = "mean",monthly_stat_per = "range") |> 
      easyrgee:::ic_add_month_property() # adds moy property for join later
    
    moi_sd_composites <- easyrgee::ic_yr_mo_composite_stat(ic = ic,
                                      month_range = c(mr[1],mr[2]),
                                      year_range = c(baseline_years[1],baseline_years[2]),
                                      stat = "sd",monthly_stat_per = "range") |> 
      easyrgee:::ic_add_month_property() # adds moy property for join later
      
    # a little hacky fix in v0.2
    assessment_yr<- lubridate::year(last_img_meta$time_start)
    current_yr_mo_composites <-  easyrgee::ic_yr_mo_composite_stat(ic = ic,
                                                                   month_range = c(mr[1],mr[2]),
                                                                   year_range = c(assessment_yr,assessment_yr),
                                                                   stat="mean",
                                                                   monthly_stat_per = "range"
                                                                        )$map(
                                                                          function(img){
                                                                            img$select("NDVI_mean")$rename("NDVI") 
                                                                            
                                                                            }
                                                                          ) 
      current_yr_mo_composites <- current_yr_mo_composites |> easyrgee:::ic_add_month_property() # adds moy property for join later
      
    # current_yr_mo_composites |> ee_get_date_ic()
    # moi_mean_composites|> ee_get_date_ic()
    # moi_sd_composites |> ee_get_date_ic()
    # new func
    #export ic_join_bands
 
 current_and_baseline_stats <-  current_yr_mo_composites |> 
   easyrgee:::ic_join_bands(y = moi_mean_composites,by = "moy") |>  
   easyrgee:::ic_join_bands(y=moi_sd_composites,by = "moy") 
  
    # current_and_baseline_stats |> ee_get_date_ic()
    
    
    # this is still a bit ugly - opened issue wrap simplifier
    ndvi_zscore<- current_and_baseline_stats$map(
      function(img){
        zscore<- img$expression(
          "float((NDVI-NDVI_mean)/(NDVI_stdDev))",
          opt_map= list(NDVI= img$select("NDVI"),
                        NDVI_mean= img$select("NDVI_mean"),
                        NDVI_stdDev= img$select("NDVI_stdDev")
      )
    )$rename("NDVI_z_score")
    img$select("NDVI")$addBands(zscore)
  }
  
)
    
    cat("extracting value to points")
    
    # does this work for points
    # only about 90s for 1569 pts
    
    # ndvi_zscore |> ee_get_date_ic()
    # ndvi_zscore |> ee_print()
    
    
    
    
    
    
    
    
    
    # sf:::st_geometry_type(dat2)
    # dat2 <-  dat |>slice(1:3) 
    # dat2 |> class()
    # n_distinct(sf::st_geometry_type(dat2))
    # ee_extract(x = ndvi_zscore |> easyrgee:::map_date_to_bandname_ic(),y= dat2["new_uid"] ,scale = 500 )
    # dat_w_NDVI |> count(date)
    
    # ndvi_zscore |> ee_print()
    # need to test times - about 100s for 1569 pts
    # dat_ck<- dat |> slice(1:3) |> sf::st_buffer(0)
    # debugonce(easyrgee::ee_extract_long)
    
     
    ###############################################################################
    # TESTING EE_EXTRACT DOESNT LIKE POINTS OR < 10 M BUFFERS
    ###############################################################################
    # dat2 <-  dat |>slice(1:3) |> sf::st_transform(crs=32646) |>sf::st_buffer(dist = 10) |> sf::st_transform(crs=4326)
    # ck_res <- rgee::ee_extract(x = ndvi_zscore,
    #                            y = dat2["new_uid"],
    #                            fun =ee$Reducer$mean(),
    #                            scale = scale,sf=T)
    #
    ############################################################################
    
    # maybe it should always be a buffer of at least 10 m.... does GEE processing slow down with many requests?
    # what about null locations? -- should probably be removed first?
    # 10 m buffer time?
    # dat10 <-  dat  |> sf::st_transform(crs=32646) |>sf::st_buffer(dist = 10) |> sf::st_transform(crs=4326)
    dat100 <-  dat  |> sf::st_transform(crs=32646) |>sf::st_buffer(dist = 100) |> sf::st_transform(crs=4326)
    
    # dat100_samp <-  dat100 |> slice(1:10)
    
    
    # how does this go in 5 seconds
    # extract100_samp <- rgee::ee_extract( x = ndvi_zscore |> easyrgee:::map_date_to_bandname_ic(),y = dat100_samp["new_uid"],scale = 250)
    
    
    
    ##############################################
    # if ee_extract_log does not work consistently I am going to have to figure out why/if wide does
    #  is there silent fail going on? 
    # extract100_full <- rgee::ee_extract( x = ndvi_zscore |> easyrgee:::map_date_to_bandname_ic(),y = dat100,scale = 250)
    
    # debugonce(easyrgee::ee_extract_long)
    # and this takes 100? something
    #################################################
    # this last run with 100 m buffers took 110s
    dat_w_NDVI<- easyrgee::ee_extract_long(ic = ndvi_zscore,
                                           sf = dat100,
                                           sf_col = "new_uid",
                                           scale = scale,reducer = "mean")
    
    
  

    cat("finished extracting points")
    dat_w_NDVI |> count(date)
    dat_w_NDVI |> 
      mutate(
        mo=lubridate::month(date),
        mo_diff= lubridate::month(last_img_month) - mo,
        # hack - fix above would be nicer
        parameter = if_else(str_detect(string = parameter,"NDVI_z"),"NDVI_Z",parameter),
        new_col_name = glue::glue("rs.{parameter}_{mo_diff}mo_prev_{src_suffix}"),
        new_col_name =if_else(mo_diff==0,str_replace_all(new_col_name,"0mo_prev","mo_curr"),as.character(new_col_name))
        
      ) |> 
      
      select(new_uid,new_col_name, value) |> 
      pivot_wider(names_from = new_col_name,values_from = value)
      
    

    
    
    
  
  
  }
  
  


```



need to check monthly and see how up to date JRC monthly is
```{r}

# https://gis.stackexchange.com/questions/404629/earth-engine-calculation-of-distance-between-points-to-nearest-water
# https://gis.stackexchange.com/questions/408804/forcing-gees-fastdistancetransform-to-operate-at-specific-scale
dat
# data()

library(rgee)
ee_Initialize()
JRC_yearly <- rgee::ee$ImageCollection("JRC/GSW1_3/YearlyHistory")
# JRC_yearly |> ee_print()
# debugonce(ee_get_date_ic)
# JRC_yearly |> ee_get_date_ic()
# JRC_yearly$getInfo()
JRC_example <- JRC_yearly$
  filterDate(ee$Date("2020-04-01")$advance(-5,"year"),"2021-04-01")

JRC_example |> ee_print()
JRC_example |> ee_get_date_ic()

thresholded <- JRC_example$min()$gt(1)$selfMask()
ee$Image(thresholded) |> ee_print()
nominalScale <- 30
minPixels <-3

# (nominalScale$pow(2)) * minPixels
# nominalScale


minArea <-  (nominalScale^2)*minPixels

ee$Image$pixelArea()$getInfo()

pixelCount <- thresholded$connectedPixelCount( maxSize= 100)
area <- pixelCount$multiply(ee$Image$pixelArea())
largeWaterBodies = thresholded$updateMask(area$gte(minArea))
largeWaterBodies |> ee_print()

distance_raster <-  largeWaterBodies$mask()$
  fastDistanceTransform({
    neighborhood=1024
  })$multiply(ee$Image$pixelArea())$sqrt()$rename("distance_to")

distance_raster |> 
  ee_print()

library(sf)
dat_pt_sample |> glimpse()
dat_pt_sample<- dat |> 
  slice(1:3) |> 
  st_as_sf(coords= c("_gps_reading_longitude"  ,"_gps_reading_latitude"  ), crs= 4326) |> 
  mutate(across(.cols = c("end_survey","electricity_grid"),.fns = ~as.character(.x)))

dat_pts_sample_ee <- sf_as_ee(dat_pt_sample |> st_geometry())
# dat_pts_sample_ee |> ee_print()
distanceFromMultiPoint = distance_raster$
  sampleRegions(
    
    collection= dat_pts_sample_ee,
    scale= nominalScale#,
    # geometries= true
    
  )

# distanceFromMultiPoint |> ee_print()
# distanceFromMultiPoint$getInfo()
distanceFromMultiPoint$aggregate_array("distance_to")$getInfo()

imageCol_masked <- JRC_example$map(
  function(image){
    image_masked = image$
      gt(1)$
      selfMask()
    
    return(ee$Image(image_masked$copyProperties(image,image$propertyNames())))
      
  }
)
imageCol_masked |> ee_get_date_ic()

JRC_example$first()$pixelArea()$getInfo()

ee$Image$pixelArea()$getInfo()

rgee::ee_extract
#' closest_distance_to_raster_value
#' @param x ee$Image or ee$ImageCollection
#' @param y ee$Geometry$*, ee$Feature, ee$FeatureCollection, sfc or sf objects.
#' @param geeFC
#' @param scale
#' @param min_pixels


sf
ck1 <- closest_distance_to_raster_feature(x = JRC_example,y=sf, scale=30, min_pixels = 1, search_radius = 100)
ck1 <- closest_distance_to_raster_feature(x = JRC_example,y=sf, scale=30, min_pixels = 3, search_radius = 100)
ck2 <- closest_distance_to_raster_feature(x = JRC_example,y=sf, scale=30, min_pixels = 3, search_radius = 5)
ck3 <- closest_distance_to_raster_feature(x = JRC_example,y=sf, scale=30, min_pixels = 50000, search_radius = 1000)
ck4 <- closest_distance_to_raster_feature(x = JRC_example,y=sf, scale=30, min_pixels = 3, search_radius = 1000)

closest_distance_to_raster_feature <- function(x=JRC_example,y ,scale=30,min_pixels=3,search_radius=100){
  min_area <-  (scale^2)*min_pixels
  
  cat(crayon::green("masking x image/imageCollection\n"))
  x_masked <- x$map(
    function(image){
      image_masked = image$
        gt(1)$
        selfMask()
      
      return(ee$Image(image_masked$copyProperties(image,image$propertyNames())))
      
    }
  )
  cat(crayon::green("Calculating connected components and applying size threshold \n"))
  x_thresholded<- x_masked$map(
    function(image){
      pixel_count = image$connectedPixelCount(search_radius)
      area = pixel_count$multiply(ee$Image$pixelArea())
      # area = pixel_count$multiply(image$pixelArea())
      x_gt_min_size = image$updateMask(area$gte(min_area))
      return(x_gt_min_size)
    }
  )
  # imageCol_size_gt_min |> ee_print()
  # imageCol_size_gt_min |> ee_get_date_ic()
  cat(crayon::green("Generating distance raster(s)\n"))
  euclidean_distance_to_x <-  x_thresholded$
    map(
      function(image){
        distances = image$mask()$
          fastDistanceTransform({
            neighborhood=1024
          })$multiply(ee$Image$pixelArea())$sqrt()$rename("distance_to")
        return(ee$Image(distances$copyProperties(image,image$propertyNames())))
        
      }
    )
  # euclidean_distance_imageCol |> ee_print()
  # euclidean_distance_imageCol |> ee_get_date_ic()
  # debugonce(exploreRGEE::ee_timeseries)
  cat(crayon::green("Extracting distance raster values to y\n"))
  res <- exploreRGEE::ee_timeseries(imageCol = euclidean_distance_to_x$select("distance_to"),geom=sf,temporal="all",scale=30)
  return(res)
}

JRC_example$first()$eq(ee$List(list(2,3))) 

# ck2 <- ee_closest_distance_to_val(x = JRC_example$first(),y=sf,val=2)
ck1 <- ee_closest_distance_to_val(x = JRC_example,y=sf,boolean=">=",val=2,scale=30)
ck2 <- ee_closest_distance_to_val(x = JRC_example$first(),y=sf,boolean=">=",val=2,scale=30)

debugonce(ee_closest_distance_to_val)
ck1$first() |> ee_print()
ck2 <- JRC_example$first()$eq(2)$selfMask()
ck2 |> ee_print()


switch_boolean <-  function(boolean,val){switch(boolean,
                                       ">" = function(x)x$gt(val),
                                       ">=" = function(x)x$gte(val),
                                       "<" = function(x)x$lt(val),
                                       "<=" = function(x)x$lte(val),
                                       "=" = function(x)x$eq(val),
                                       NULL

)
}
ee_bool<- switch_boolean(boolean = ">",val = 25)

ee_closest_distance_to_val <-  function(x,...){
  UseMethod('ee_closest_distance_to_val')
  
  
}
ee_closest_distance_to_val.ee.imagecollection.ImageCollection <-  function(x,
                                                                           y,
                                                                           boolean="=", 
                                                                           val=2,
                                                                           scale
                                                                           ){
  
  # stopifnot(!is.null(x), inherits(x, "ee.imagecollection.ImageCollection"))
  assertthat:::assert_that(!is.null(x), inherits(x, "ee.imagecollection.ImageCollection"))
  boolean_mask_cond<-switch_boolean(boolean=boolean,val=val)
  
  # if(is.numeric(val)){}
  # cat(crayon::green("masking x image/imageCollection\n"))
  x_masked <- x$map(
    function(image){
      image_masked = boolean_mask_cond(image)$ 
        selfMask()
      return(ee$Image(image_masked$copyProperties(image,image$propertyNames())))
      
    }
  )
  
    cat(crayon::green("Generating distance raster(s)\n"))
  euclidean_distance_to_x <-  x_masked$
    map(
      function(image){
        distances = image$mask()$
          fastDistanceTransform({
            neighborhood=1024
          })$multiply(ee$Image$pixelArea())$sqrt()$rename("distance_to")$
          reproject(crs="EPSG:4326",scale=scale)
        return(ee$Image(distances$copyProperties(image,image$propertyNames())))
        
      }
    )
  cat(crayon::green("Extracting distance raster values to y\n"))
  res <- exploreRGEE::ee_timeseries(imageCol = euclidean_distance_to_x$select("distance_to"),geom=sf,temporal="all",scale=30)
  return(res)
}

JRC_example$first() |> class()  
ee_closest_distance_to_val.ee.image.Image<-  function(x,
                                                      y,
                                                      boolean="=", 
                                                      val=2,
                                                      scale
){
  
  # stopifnot(!is.null(x), inherits(x, "ee.imagecollection.ImageCollection"))
  assertthat:::assert_that(!is.null(x), inherits(x, "ee.image.Image"))
  boolean_mask_cond<-switch_boolean(boolean=boolean,val=val)
  
  # if(is.numeric(val)){}
  cat(crayon::green("masking x image/imageCollection\n"))
  x_masked <-ee$Image(
    boolean_mask_cond(x)$
      selfMask()$
      copyProperties(x,x$propertyNames())
    )
  cat(crayon::green("Generating distance raster(s)\n"))
  
        FDT = x_masked$mask()$
          fastDistanceTransform({
            neighborhood=1024
          })
        
        distances= FDT$
          multiply(ee$Image$pixelArea())$
          sqrt()$
          rename("distance_to")$
          reproject(crs="EPSG:4326",scale=scale)
        
  euclidean_distance_to_x <-ee$Image(distances$copyProperties(x_masked,x_masked$propertyNames()))
  
  cat(crayon::green("Extracting distance raster values to y\n"))
  res <- exploreRGEE::ee_timeseries(imageCol = euclidean_distance_to_x$select("distance_to"),geom=sf,temporal="all",scale=30)
  return(res)
}
  
  
thresholded <- JRC_example$min()$gt(1)$selfMask()
ee$Image(thresholded) |> ee_print()
nominalScale <- 30
minPixels <-3

# (nominalScale$pow(2)) * minPixels
# nominalScale




pixelCount <- thresholded$connectedPixelCount( maxSize= 100)
area <- pixelCount$multiply(ee$Image$pixelArea())
largeWaterBodies = thresholded$updateMask(area$gte(min_area))
largeWaterBodies |> ee_print()

distance_raster <-  largeWaterBodies$mask()$
  fastDistanceTransform({
    neighborhood=1024
  })$multiply(ee$Image$pixelArea())$sqrt()$rename("distance_to")

distanceFromMultiPoint = distance_raster$
  sampleRegions(
    
    collection= dat_pts_sample_ee,
    scale= scale#,
    # geometries= true
    
  )

# distanceFromMultiPoint |> ee_print()
# distanceFromMultiPoint$getInfo()
distanceFromMultiPoint$aggregate_array("distance_to")$getInfo()



  
}




```


extract_pixel_values (inspired by geemap). Could be useful for general work as well as developing the `Inspector` feature

```{r}
# https://geemap.org/common/

#' Extract pixel values of `ee$Image` or `ee$ImageCollection`
#' @description Samples the pixels of an image, returning them as a ee.Dictionary.
#' @param dat sf object
#' @param ee_object ee$Image or ee$ImageCollection
# data(dat)
# dat |> slice(1:3)
# rgee::ee_extract()
library(sf)
library(rgee)
rgee::ee_Initialize()
dat <-  read_csv("data/example_raw_data.csv")
dat <- dat |> slice(1:3)
sf <- st_as_sf(dat,coords=c("_gps_reading_longitude","_gps_reading_latitude"),crs=4326) |> 
  select(`_uuid`)

sf_buffer <-  sf |> 
  st_transform(crs=32646) |> 
  st_buffer(dist= 10) |> 
  st_transform(crs=4326)
JRC_yearly <- rgee::ee$ImageCollection("JRC/GSW1_3/YearlyHistory")
# JRC_yearly |> ee_print()
JRC_yearly$aggregate_array("system:time_start")$getInfo()

# debugonce(ee_get_date_ic)
# JRC_yearly |> ee_get_date_ic()
# JRC_yearly$getInfo()
JRC_example <- JRC_yearly$
  filterDate(ee$Date("2020-04-01")$advance(-5,"year"),"2021-04-01")

JRC_example |> ee_get_date_ic()

extract_pixel_values <- function(sf,ee_object=JRC_example, scale=30){
   
  cat("uploading sf feature as temporary cloud asset")
  # assert_that (no_logical columns in sf)
  fc |> ee_print()
  fc <- sf_as_ee(sf)
  
  # try on modis see if it works
  bbox_ee<- rgee::sf_as_ee(sf::st_bbox(sf) |> sf::st_as_sfc())
  ee_object <- rgee::ee$ImageCollection(ic_url)$filterBounds(bbox_ee)
  ee_object <-  ee_object$filterDate("2021-01-01","2021-04-01")
  
  ee_object_image = ee_object$first()
  ee_object_image |> class()
  ee_object_first <- ee_object$first()
  ee_object |> ee_get_date_ic()
  # ee_object_first |> ee_print()
  debugonce(exploreRGEE:::extract_time)
  ck1 <- exploreRGEE::ee_timeseries(imageCol = ee_object$select("NDVI"),geom=sf,temporal="all",scale=30)
  exploreRGEE::ee_timeseries(imageCol = ee_object,geom=sf_buffer,temporal="all",scale=30)
  # exploreRGEE::ee_timeseries(imageCol = ee_object_first,geom=sf_buffer,temporal="all",scale=30)
  
  fc_pixel_vals <- ee_object_first$sampleRegions( 
       collection= fc,
       scale= scale
       #,
    # geometries= true
    
  )
  # ee_object$sampleRegions
  res$getInfo()
  res <- fc_pixel_vals
  res$get("waterClass")
  res$aggregate_array("waterClass")$getInfo()
  res |> ee_print()
  
  res$getInfo() |> ee_print()
  ee_object |> ee_print()
  
  fc_pixel_vals <- ee_object$map(
    function(img){
      img$sampleRegions(
       collection= fc,
       scale= scale
       #,
    # geometries= true
    
  )
    }
  )
      
  
  
  
}
```

- parameters that make sense when aggregated:
  + maybe start of rainy season date?



```{r}
# var getCloudScores = function(img){
#     //Get the cloud cover
#     var value = ee.Image(img).get('CLOUD_COVER');
#     return ee.Feature(null, {'score': value})
# };
# 
# var results = landsat.map(getCloudScores);
# print(Chart.feature.byFeature(results));

#' get cloud scores
#' @param img image
#' @return data.frame containing image date and cloud score
#' 


getCloudScores <-  function(img){
  value <-  ee$Image(img)$get("CLOUD_COVER")
  
}

```


```{r}

# # do we want NDVI of closest image or median composite say 1 month prior?
# closest_img_date <-  function(ic, date){
#   
#   date <-  lubridate::ymd(date)
#   ee_date <- rgee::ee$Date(date)
#   
#   
# }
# 
# ic = ic.map(function(image){
#   return image.set(
#     'dateDist',
#     ee.Number(image.get('system:time_start')).subtract(dateOfInterest.millis()).abs()
#   );
# });
# This new field 'dateDist' can now be used to sort the image collection.
# 
# ic = ic.sort('dateDist');

```






```{r}

# from geemap for points i think
dict_values_tmp = ee_object.reduceRegion(
                            reducer=self.roi_reducer,
                            geometry=self.user_roi,
                            scale=scale,
                            bestEffort=True,
                        ).getInfo()


```

