---
title: "Agricultural Event Detection"
output: html_document
bibliography: references.bib
link-citation: true
csl: remote_sensing_journal.csl
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Intro

Using this document to brainstorm ideas on agricultural event detection.

Agricultural event of interest include:

- tilling/plowing/mowing
- early phase declines?

## Methods

Most methods tend to use sentinel-2 or a combination of sentinel-1 & sentinel-2. Most methods seem to rely on a time series analysis of a few key RS indicators including:

- NDVI (sentinel-2)
- Bare soil marker [@sentinelhub_bare_soil_marker_sep2020]
- Backscatter (sentinel-2)[@sentinel_hub_area_monitoring_service]


## Challenge

 bladadafdsaf  [@sinergize]
While creating RS indicators and performing time-series analysis would be relatively straight forward using standard GEE tools, there are two areas which will require further research.

1. crop-land delineation/segmentation
2. asdfdsafa


## Resources
- [Separability of Mowing and Ploughing Events on Short Temporal Baseline Sentinel-1 Coherence Time Series](https://www.mdpi.com/2072-4292/12/22/3784)
- [Area Monitoring — Mowing Marker](https://medium.com/sentinel-hub/area-monitoring-mowing-marker-e99cff0c2d08)
- [Mapping fallow fields using Sentinel-1 and Sentinel-2 archives over farming-pastoral ecotone of Northern China with Google Earth Engine](https://www.tandfonline.com/doi/full/10.1080/15481603.2022.2026638)
- [Area Monitoring — Bare Soil Marker](https://medium.com/sentinel-hub/area-monitoring-bare-soil-marker-608bc95712ae)


```{r setup}
library(surveyGEER)
```


# References

