---
title: "ja-understanding_modis_cloud_masking"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ja-understanding_modis_cloud_masking}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(sf)
library(tidyrgee)
library(targets)
devtools::load_all()
tar_load()

# create center point for map
map_center_pt_ee <- sf_as_ee(sf::st_bbox(col_pt_data_clean) |> 
                               sf::st_as_sfc() |> 
                               st_centroid()
)
```

I want to know what happens when pixels with masked values are composited.
Therefore lets, do some modis cloud masking which will pixels
```{r}
modis_drought_ic <- get_modis_drought_indicators()
modis_drought_ic_cloud <- get_modis_drought_indicators(mask = "cloud")
modis_drought_ic_cloud_quality <- get_modis_drought_indicators(mask = "cloud&quality")
```


Then let's look at masked NDVI for april 2022, may 2022, and may-april 2022 composite
- first get the images
```{r}
# may_modis <- modis_drought_ic |> 
#   filter(year==2022,month==5) |> 
#   as_ee()
april_modis_cloud <- modis_drought_ic_cloud |> 
  filter(year==2022, month== 4) |> 
  as_ee()
may_modis_cloud <- modis_drought_ic_cloud |> 
  filter(year==2022, month== 5) |> 
  as_ee()

april_may_composite_cloud <-  modis_drought_ic_cloud |> 
  filter(year==2022, month%in% c(4,5)) |> 
  summarise(stat="mean")
```

- then visualize
  + so looks like when the data is masked in 1 image, but not in the other it returns the unmasked value. This is good and as expected, but was not sure.

```{r}

Map$centerObject(map_center_pt_ee,zoom=7)
m1 <- Map$addLayer(april_modis_cloud,
             visParams = list(min=0,
                              max=1,
                              palette=c("red","yellow","green"),
                              bands="NDVI"),
             name = "april cloud masked")
m2 <- Map$addLayer(may_modis_cloud,
             visParams = list(min=0,
                              max=1,
                              palette=c("red","yellow","green"),
                              bands="NDVI"),
             name = "may cloud masked")

m3 <- Map$addLayer(april_may_composite_cloud$ee_ob,
             visParams = list(min=0,
                              max=1,
                              palette=c("red","yellow","green"),
                              bands="NDVI_mean"),
             name = "april-may cloud masked composite")


m4 <- m1+m2+m3
leafgl::addGlPoints(map=m4,data = col_pt_data_clean)

# may_modis_cloud_quality <- modis_drought_ic_cloud_quality |> 
#   filter(year==2022, month== 5) |> 
#   as_ee()
```

## How to deal with clouds
- by using cloud + quality mask we can reduce the size of the mask
- however, there will still be many masked points...
- gap filling
  + run a focal window (kernel?) on the masked image, we should be able to take a neighborhood mean using only unmasked pixels? 
  + use the last non-cloudy pixel.... but we shouldn't allow it to go too far back? maybe one month is acceptable
  + The Terra and Aqua combination (TAC) method is the simplest and most transparent of the temporal methods [@li2020gap] , [@parajka2008modistac]
  + Interpolation with regression (https://groups.google.com/g/google-earth-engine-developers/c/SLUrrjZqfDQ/m/dCT0FiyTAAAJ)
  (https://code.earthengine.google.com/d66813de389e3733922419d342e08c83)
  https://spatialthoughts.com/2021/11/08/temporal-interpolation-gee/#more-6094
### try TAC method
-i've adjuste the `get_modis_drought_indicators` to allow either satellite
```{r}

# terra
# terra <- ee$ImageCollection("MODIS/061/MOD13Q1")
# aqua - check check
# aqua <- ee.ImageCollection("MODIS/061/MYD13Q1")
terra_drought_cloud <- get_modis_drought_indicators(mask = "cloud",satellite = "terra")
aqua_drought_cloud <- get_modis_drought_indicators(mask = "cloud",satellite = "aqua")

terra_may_cloud <- terra_drought_cloud |> 
  filter(year==2022, month== 4) |> 
  as_ee()

aqua_may_cloud <- aqua_drought_cloud |> 
  filter(year==2022, month== 4) |> 
  as_ee()


# Map$centerObject(map_center_pt_ee,zoom=7)
map_terra_may_cloud <- Map$addLayer(terra_may_cloud,
             visParams = list(min=0,
                              max=1,
                              palette=c("red","yellow","green"),
                              bands="NDVI"),
             name = "terra may - cloud masked")

map_aqua_may_cloud <- Map$addLayer(aqua_may_cloud,
             visParams = list(min=0,
                              max=1,
                              palette=c("red","yellow","green"),
                              bands="NDVI"),
             name = "aqua may - cloud masked")

map_terra_aqua_may_cloud <- map_terra_may_cloud+map_aqua_may_cloud
leafgl::addGlPoints(map=map_terra_aqua_may_cloud,data = col_pt_data_clean)
```

- TAC is rudimentarily impleented and does increase % non-clouds
- by taking the `mean` it fills gaps... not sure if it is always right to take the mean
```{r}
terra_aqua <- bind_ics(list(terra_drought_cloud,aqua_drought_cloud))

terra_aqua_composite <- terra_aqua |> 
  group_by(year,month) |> 
  summarise(stat="mean")

may_terra_aqua_composite <- terra_aqua_composite |> 
  filter(year==2022, month==4)
map_terra_aqua_cloud_composite_may<-  Map$addLayer(may_terra_aqua_composite$ee_ob,
             visParams = list(min=0,
                              max=1,
                              palette=c("red","yellow","green"),
                              bands="NDVI_mean"),
             name = "terra-aqua cloud masked composite")

map_terra_aqua_full <- map_terra_aqua_may_cloud+map_terra_aqua_cloud_composite_may
leafgl::addGlPoints(map=map_terra_aqua_full,data = col_pt_data_clean)
```


lets try interpolation of the TAC-GAP filled IC
```{r}
interp_tac <- interpolate_ic(x = terra_aqua_composite$ee_ob,time_window = 40 )
```




now lets visualize - i changed month to 4 here and above because we can't interpolate on last month data

```{r}
interp_tac_tidy <- as_tidyee(interp_tac)
interp_tac_may <- interp_tac_tidy |> 
  filter(year==2022, month ==4) |> 
  as_ee()

map_with_interpolated_ndvi<-  Map$addLayer(interp_tac_may,
             visParams = list(min=0,
                              max=1,
                              palette=c("red","yellow","green"),
                              bands="NDVI_mean"),
             name = "Interpolated terra-aqua cloud masked composite")
another_map <-  map_terra_aqua_full+map_with_interpolated_ndvi
leafgl::addGlPoints(map=another_map,data = col_pt_data_clean)
```

# Terra Vs Aqua....
I want to understand how different `NDVI` values are from Terra vs Aqua satellites
```{r}

terra_aqua_ndvi_joined <- terra_drought_cloud |> 
  select(terra_ndvi="NDVI") |> 
  inner_join(
    aqua_drought_cloud |> 
      select(aqua_ndvi="NDVI")  , by ="month"
  )
april_terra_aqua <- terra_aqua_ndvi_joined |> 
  filter(year==2021,month == 4)
april_terra_aqua <- ee$Image(terra_aqua_ndvi_joined$ee_ob$first())

# tar_load_everything()
rnd_sample <- sf_as_ee(sf::st_bbox(col_pt_data_clean) |> 
                               sf::st_as_sfc() |> 
           st_sample(size = 1000)
)

rnd_sample_vals <- april_terra_aqua$sampleRegions(
  collection= rnd_sample,
  scale = 250
)
rnd_sample_vals_df <- ee_as_sf(rnd_sample_vals)
rnd_sample_vals_df |> 
  mutate(difference = terra_ndvi - aqua_ndvi) |> 
  summarise(
    avg_diff=mean(difference,na.rm=T)
  )

rnd_sample_vals_df |> 
  ggplot(aes(x= terra_ndvi, y= aqua_ndvi))+
  geom_point()
  



select_aqua_terra <- function(img){
  aqua <- img$select("aqua_ndvi")
  terra <- img$select("terra_ndvi")
  ee$Image(-999)$
    where(terra$eq(0),aqua)$
    where(terra$neq(0),terra)
}
ck2 <- select_aqua_terra(april_terra_aqua)


ck_map <- Map$addLayer(ck2$updateMask(ck2$neq(-999)),
             visParams = list(min=0,
                              max=1,
                              palette=c("red","yellow","green"),
                              bands="constant"),
             name = "ck")
all_maps <- ck_map+another_map
leafgl::addGlPoints(map=all_maps,data = col_pt_data_clean)


ck3 <- april_terra_aqua$select("terra_ndvi")$
  unmask(april_terra_aqua$select("aqua_ndvi"),sameFootprint=TRUE)
ck3$bandNames()$getInfo()
comparison_map <- Map$addLayer(
  # ck2$updateMask(ck2$eq(-999))
  ck3
  ,
             visParams = list(min=0,
                              max=1,
                              palette=c("red","yellow","green"),
                              bands="terra_ndvi"
                              # bands="constant"
                              
                              ),
             name = "ck")+
  Map$addLayer(april_terra_aqua,
                visParams = list(min=0,
                              max=1,
                              palette=c("red","yellow","green"),
                              bands="terra_ndvi"), name = "terra")+
  Map$addLayer(april_terra_aqua,
                visParams = list(min=0,
                              max=1,
                              palette=c("red","yellow","green"),
                              bands="aqua_ndvi"), name = "aqua")
```


the function has been updated to include the lessons of the above experiments.
1. cloud mask alone is too aggressive
2. 
```{r}

# Aggressive Cloud Masking
ndvi_cmask_tac <- get_modis_drought_indicators(mask = "cloud",
                             satellite = "terra",
                             TAC = T,
                             temporal_interpolation = F)

ndvi_cmask_tac_interp <- get_modis_drought_indicators(mask = "cloud",
                             satellite = "terra",
                             TAC = T,
                             temporal_interpolation = T)

ndvi_cmask <- get_modis_drought_indicators(mask = "cloud",
                             satellite = "terra",
                             TAC = F,
                             temporal_interpolation = F)

cmask_comparisons<- dplyr::lst(ndvi_cmask, ndvi_cmask_tac,ndvi_cmask_tac_interp)

cmask_maps <- cmask_comparisons |> 
  purrr::map2(.y=names(cmask_comparisons),.f = ~{
    img_filt_ee <- .x |> 
              filter(year==2022, month == 4) |>
      as_ee()
  Map$addLayer(img_filt_ee,visParams = list(min=0, max=0.7, palette=c("red","yellow","green","darkgreen"),bands="NDVI"),name=.y)
              }
  )
            
cmask_maps$ndvi_cmask+cmask_maps$ndvi_cmask_tac+cmask_maps$ndvi_cmask_tac_interp


# Less Aggressive
ndvi_cqmask_tac <- get_modis_drought_indicators(mask = "cloud&quality",
                             satellite = "terra",
                             TAC = T,
                             temporal_interpolation = F)

ndvi_cqmask_tac_interp <- get_modis_drought_indicators(mask = "cloud&quality",
                             satellite = "terra",
                             TAC = T,
                             temporal_interpolation = T)

ndvi_cqmask <- get_modis_drought_indicators(mask = "cloud&quality",
                             satellite = "terra",
                             TAC = F,
                             temporal_interpolation = F)

cqmask_comparisons<- dplyr::lst(ndvi_cqmask, ndvi_cqmask_tac,ndvi_cqmask_tac_interp)

cqmask_maps <- cqmask_comparisons |> 
  purrr::map2(.y=names(cmask_comparisons),.f = ~{
    img_filt_ee <- .x |> 
              filter(year==2022, month == 4) |>
      as_ee()
  Map$addLayer(img_filt_ee,visParams = list(min=0, max=0.7, palette=c("red","yellow","green","darkgreen"),bands="NDVI"),name=.y)
              }
  )
            
cqmask_maps$ndvi_cqmask+
  cqmask_maps$ndvi_cqmask_tac+
  cqmask_maps$ndvi_cqmask_tac_interp



comparison_map2 <- comparison_map+
   Map$addLayer(ndvi_cqmask_tac_interp |> filter(year==2022,month==4) |> as_ee(),
                visParams = list(min=0,
                              max=1,
                              palette=c("red","yellow","green"),
                              bands="NDVI"), name = "ndvi_cqmask_tac_interp")

leafgl::addGlPoints(map=comparison_map2,data = col_pt_data_clean)
```
