---
title: "ka-irq_water_classification"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ka-irq_water_classification}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(surveyGEER)
library(rgee)
library(tidverse)
library(sf)
library(tidyrgee)
devtools::load_all()
ee_Initialize()
```


Load in AOI and check on map
```{r}
geom <- ee$Geometry$Polygon(list(
      c(44.2354847930793, 34.83077069846819),
          c(44.261577322376176, 34.692001577255105),
          c(44.40851946104805, 34.511140220037795),
          c(44.607646658313676, 34.47152432533435),
          c(44.687297537219926, 34.58918498051208),
          c(44.5211293243293, 34.75069665768057),
          c(44.393413259876176, 34.810477595189816),
          c(44.28217668761055, 34.88598770009403),
          c(44.223125173938676, 34.84316957807562)
    ))
Map$addLayer(geom,{},"geom")
```


**Some pre-processing**
- filter temporally and spatially to area of interest
- remove images with >= 25% cloud cover
- rename bands

```{r}

l8_ic <-  ee$ImageCollection('LANDSAT/LC08/C02/T1_L2')$
  filterDate("2013-01-01","2022-12-31")$
  filterBounds(geom)$
  filter(ee$Filter$lt('CLOUD_COVER', 25))$
  select(list('SR_B2', 'SR_B3', 'SR_B4','SR_B5', 'SR_B6', 'SR_B7'),
         list('blue', 'green', 'red', 'nir', 'swir1', 'swir2')
  )
```

- Convert to tidy format for easy monthly composites
- Filter to just April months
```{r}
l8_ic_tidy <-   as_tidyee(l8_ic)


l8_median_april <- l8_ic_tidy |> 
  group_by(year, month) |> 
  summarise(
    stat="median"
  ) |> 
  ungroup() |> 
  filter(month==4) 
```


- Create MNDWI Masks
```{r}

l8_median_april_mndwi1 <- l8_median_april$ee_ob$map(
  function(img){
    time_start <- img$get("system:time_start")
    img_mndwi <- img$normalizedDifference(list('green_median', 'swir1_median'))$
      rename("mndw1")
      return(img_mndwi$set("system:time_start",time_start))
  }
  
)

l8_median_april_mndwi2 <- l8_median_april$ee_ob$map(
  function(img){
    time_start <- img$get("system:time_start")
    img_mndwi <- img$normalizedDifference(list('green_median', 'swir2_median'))$
      rename("mndw2")
    return(img_mndwi$set("system:time_start",time_start))
  }
)
```

- Convert back to tidyee class for easy joinning
- Now we have an IC with 2 bands
- we can create out water_mask using the two bands.

```{r}
mndwi_tidy <- as_tidyee(l8_median_april_mndwi1)
mndwi2_tidy <- as_tidyee(l8_median_april_mndwi2)

mndwi_tidy <- mndwi_tidy |> 
  inner_join(mndwi2_tidy, by ="system:time_start")

water_yearly = mndw_tidy$ee_ob$map(
  function(img){
    time_start <- img$get("system:time_start")
    img$
      select("mndw1")$gte(0)$
      And(img$select("mndw2")$gte(0))$
      set("system:time_start",time_start)$rename("water")
  }
)
water_yearly_tidy <- as_tidyee(water_yearly)

```

- Visualize to make sure results make sense
```{r}
years_to_map <-  2013:2022
years_to_map <- years_to_map |> 
  set_names(paste0(years_to_map,"_water"))

yearly_water_maps <- years_to_map |> 
  purrr::map2(.y =names(years_to_map),
    ~Map$addLayer(
      water_yearly_tidy |>
      filter(year==.x) |> 
      as_ee(),
      visParams= list(min=0,max=1,palette=c("white","blue")),
      name=.y
  )
  )

Map$centerObject(geom, 10)

yearly_water_maps$`2013_water`+
  yearly_water_maps$`2014_water`+
  yearly_water_maps$`2015_water`+
  yearly_water_maps$`2016_water`+
  yearly_water_maps$`2017_water`+
  yearly_water_maps$`2018_water`+
  yearly_water_maps$`2019_water`+
  yearly_water_maps$`2020_water`+
  yearly_water_maps$`2021_water`+
  yearly_water_maps$`2022_water`+
  Map$addLayer(geom,{})
```


- Extract time series
```{r}
water_pixels_in_geom <- water_yearly_tidy |> 
  ee_extract_tidy(y = geom, scale = 30,stat = "sum")
```

- Plot time series
```{r}
water_pixels_in_geom |> 
  mutate(area_km2 = (value*30*30)*1e-6) |> 
  ggplot(aes(x= date,y=area_km2))+
  # geom_point()+
  geom_line()+
  geom_point()+
  scale_x_date(date_breaks = "1 year",date_labels = "%Y")+
  labs(x="Date", y="Area (Km2)",title= "Approximate Reservoir Area")+
  theme_bw()
```
