---
title: "workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}

devtools::load_all()
library(tidyverse)
library(easyrgee)
library(surveyGEER) # loads example datset as dat
library(here)


```

## DEMO
- read in original raw data 

```{r}

dat |> glimpse()
```

- run `setup_geo_data` function to anonymize data 
- creates a `data_share` folder with anonymized coordinates 
- creates a `vault` folder with a lookup table

```{r}


setup_geo_data(df = dat,
               uuid = "_uuid",
               lon = "_gps_reading_longitude" ,
               lat = "_gps_reading_latitude" 
               )

# done
```

Let's look at the outputs - first the anonymized coordinates
The geodata is anonymous.
- this file can be sent to RS analyst for RS extraction
- the data keeper can should delete this folder after sending and keep only the vault folder

```{r}
geodat <- read_rds(file = "data_share/coords_anonymized.rds")
```


lockup is also is anonymous.
```{r}
lookup<- read_rds("vault/lookup.rds")
```

at this point it is easy to just use an application like `7zip` to make a password encrypted compressed folder for the `vault` folder.



### Encryption in R - not necessary

Below i am just playing around with the `cyphr` package. I don't think this is very necessary.


- encrypt the file
- with ssh key we are using assymetric encryption (public + private key)
```{r}
user_key_path <- cyphr::ssh_keygen(path = "vault",
                                   password = rstudioapi::askForPassword()
                                   )

cyphr::data_admin_init("vault", path_user = user_key_path)
key <- cyphr::data_key("vault", path_user = user_key_path)

cyphr::encrypt(saveRDS(lookup, "vault/lookup_encrypt.rds"), key)
readRDS("vault/lookup_encrypt.rds")

```

delete un-encrypted..... eventually we need to link new attributes

```{r}

# doesn't work
readRDS("vault/lookup_encrypt.rds")

key <- cyphr::data_key(path_data = "vault",path_user = "vault")

cyphr::decrypt(readRDS("vault/lookup_encrypt.rds"),key)

```

- cant i just use symmetric encryption since it's not meant for communication or sharing


```{r}

k <-  openssl::aes_keygen()
actual_key <- cyphr::key_openssl(key = k,mode = "cbc")
# dir.create("encrypt_sandbox")
openssl::aes_cbc_encrypt(data = "vault/lookup.rds",key = rstudioapi::askForPassword())
cyphr:::openssl_find_key()
path <- Sys.getenv("USER_KEY", "~/.ssh/id_rsa")

?Sys.getenv()
# key <- cyphr::key_openssl(
# cyphr::encrypt_data()
```







